Notes on conformal manifolds

### 2025-10-21 Draft_MLFlatDirections.tex
Small updates in the text, tildas in the figures.

{\Large Machine Learning the Conformal Manifold of Holographic CFT2’s}

@@

\begin{abstract}
	We investigate the structure of conformal manifolds around AdS$_3 \times S^3$ which lift from continuous flat directions in the scalar potential of gauged supergravity resulting from six-dimensional $\mathcal{N}=(1,1)$.
	Our approach combines numerical exploration and symbolic inference. For the latter, we develop a symbolic regression algorithm based on Annealed Sequential Monte Carlo samples, a combination of Annealed Importance Sampling and Sequential Monte Carlo samplers, well-suited to uncovering polynomial constraints in high-dimensional parameter spaces. 
	The algorithm reconstructs a set of polynomial relations that provides an explicit analytic parametrization of a new family of solutions. 
\end{abstract}

@@

Machine learning strategies have previously been applied to identify new isolated vacua in $\mathrm{SO}(8)$ supergravity \cite{Comsa:2019rcz,Berman:2022jqn}. More broadly, there has been increasing interest in applying machine learning and numerical techniques across various domains of high-energy physics. This includes, for instance, the characterisation of Calabi–Yau metrics and hypersurfaces \cite{Ashmore:2019wzb,Berman:2021mcw,Larfors:2022nep,Berglund:2022gvm,Jejjala:2020wcc,Douglas:2006rr,Larfors:2021pbb,He:2018jtw}, as well as broader efforts to explore the string theory landscape \cite{He:2017aed,Carifio:2017bov,Ruehle:2017mzq}. Additional applications include studies in CFT \cite{Chen:2020dxg}, investigations of the supergravity landscape \cite{Brady:2025zzi,Krishnan:2020sfg}, and explorations of the AdS/CFT correspondence \cite{Hashimoto:2018ftp}. More generally, machine learning has found utility in the study of string theory, geometry, and fundamental physics \cite{Ruehle:2020jrk,He:2023csq,Bao:2021auj}.

In this work, we demonstrate the viability of the aforementioned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of six-dimensional non-chiral $\mathcal{N} = (1,1)$ supergravity on ${\rm AdS}_3 \times S^3$, or to type IIB supergravity on ${\rm AdS}_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the conformal manifold with a symbolic regression technique. There exists a large body of literature on symbolic regression, using methods from genetic programming \cite{koza1994genetic,virgolin2021improving,randall2022bingo,burlacu2019parsimony}, to deep learning \cite{petersen2019deep,kamienny2022end}, generative models \cite{valipour2021symbolicgpt}, diffusion models \cite{bastiani2025diffusion}, and equation learning with the nodes being symbolic operations \cite{2018arXiv180607259S}. A state-of-the-art algorithm is AIFeynman \cite{Udrescu:2019mnk}. It uses neural networks to identify structures in the dataset (such as translational symmetries, multiplicative separability, compositionality\dots) to recursively define simpler problems on which they can fit the solutions with polynomials and basic functions.  However, this method is slow, usually limited to low-dimensional spaces, and, by construction, can only fit one expression at a time. Alternatively, we develop here a symbolic regression algorithm based on Annealed Sequential Monte Carlo sampler~\cite{zimmermann2021nested,syed2024optimisedannealedsequentialmonte} (combining the Annealed Importance Sampling~\cite{neal1998annealedimportancesampling} approach with a Sequential Monte Carlo sampler (SMC)~\cite{del2006sequential}), since this approach is much better suited to the study of geometric loci defined by intersecting polynomials in high-dimensional spaces.

@@

%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
\section{Supergravity Setup} \label{sec:sugra}
Three-dimensional $\mathcal{N}=8$ (half-maximal) gauged supergravity is governed by the Lagrangian~\cite{Nicolai:2001ac,deWit:2003ja}
%
\begin{equation}	\label{eq: lagrangian_rephrased} 
	e^{-1}\mathcal{L}=R+\frac1{8}g^{\mu\nu}D_\mu M^{\bM\bN}D_\nu M_{\bM\bN}+e^{-1}\mathcal{L}_{\text{CS}}-V,
\end{equation}
%
which comprises an Einstein-Hilbert term $R$, a kinetic term for scalar fields parametrised by the matrix $M_{\bM\bN}$, a Chern-Simons contribution $\mathcal{L}_{\text{CS}}$, and a scalar potential $V$, see app.~\ref{app:sugra} for details on the definition of the terms. The scalar degrees of freedom parametrise the coset space
%
\begin{equation}	\label{eq: scalarcoset_rephrased}
	\frac{\text{SO}(8,4)}{\text{SO}(8)\times\text{SO}(4)},
\end{equation}
%
through the symmetric matrix $M_{\bK\bL}$.\footnote{In full generality, the coset space of half-maximal supergravity in three dimensions is $\text{SO}(4+p,4+q)/\big(\text{SO}(4+p)\times\text{SO}(4+q)\big)$ for $p$ vectors and $q$ tensors in 6$d$. We consider here the case $p=4$ only.} With appropriate gauging (see eq.~\eqref{eq:embedtens}), this $\text{SO}(8,4)$ theory is a consistent truncation of six-dimensional ${\cal N}=(1,1)$ supergravity on $S^{3}$~\cite{Cvetic:2000dm,Deger:2014ofa,Hohm:2017wtr}. It features an ${\rm AdS}_{3}$ stationary point at the scalar origin ($M_{\bK\bL}=\delta_{\bK\bL}$, the ${\rm SO}(8,4)$ identity matrix), corresponding to an ${\rm AdS}_{3}\times S^{3}$ solution in 6d. In the following, we will be interested in exploring flat directions of the potential $V$ around this point, \textit{i.e.} stationary points that are continuously connected to the origin. In three dimensions, the flat directions constitute a family of ${\rm AdS}_{3}$ vacua sharing the same cosmological constant. The corresponding solutions in six dimensions are of the form ${\rm AdS}_{3}\times M^{3}$, with $M^{3}$ some deformation of the round sphere $S^{3}$, parametrised by the constant scalar vevs defining the 3d vacua.

@@

	To address these challenges, we formulate the search for annihilating polynomials as a determination of a probability density on the space of polynomials that stresses higher probabilities on polynomials $p$ that minimise the error
	\begin{equation}
		\sum_{i=1}^{n_{\rm points}}p\big(\vec{x}^{(i)}\big)^{2}.
	\end{equation}
	Polynomials sampled from this density would thus be good approximations to annihilating polynomials. Such a probability density is \textit{a priori} hard to find, but can be approximated using algorithms based on Sequential Important Sampling~\cite{Doucet2001,Liu2004} such as Sequential Monte Carlo samplers~\cite{del2006sequential}. They are inspired from Important Sampling (IS)~\cite{osti_4441935,RobertCasella2004}, a method to parametrise expectation values with respect to a target distribution $\pi$, that can be evaluated pointwise, but from which one does not know how to draw samples. IS uses another, typically simpler, distribution $\eta$ (called importance distribution) that one knows how to draw from. It is based on the observation that the expectation value
	\begin{equation}
		\mathbb{E}_{\pi}[f] = \int f(z)\pi(z)\ \d z
	\end{equation}
	for some function $f$ with respect to $\pi$ can be computed from the expectation $\mathbb{E}_{\eta}[w f]$, now computed with respect to the distribution $\eta$, with unnormalised weights $w = \pi/\eta$. Therefore, one can estimate $\mathbb{E}_{\pi}[f]$ from a Monte Carlo method: we draw a sample $\{z^{(k)}\}_{k\in\llbracket 1, n_{\rm particles}\rrbracket}$ from $\eta$ and use the weighted average
	\begin{equation}
		\mathbb{E}_{\pi}[f] \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} w\big(z^{(k)}\big)\,f\big(z^{(k)}\big)\Bigg/\displaystyle \sum_{i=1}^{n_{\rm particles}} w\big(z^{(k)}\big).
	\end{equation}
	Stated differently, the target distribution $\pi$ can be approximated from a weighted collection of particles $\{z^{(k)},w^{(k)}\}_{k\in\llbracket 1, n_{\rm particles}\rrbracket}$ drawn from the importance distribution $\eta$ :
	\begin{equation}
		\pi(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}\,\delta\big(z-z^{(k)}\big)\Bigg/\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)},
	\end{equation}
	where $w^{(k)} = w\big(z^{(k)}\big)$. The sample admits $\pi$ as marginal distribution asymptotically when $n_{\rm particles}\rightarrow+\infty$. The accuracy of this estimation depends on the size of the sample and, crucially, on the choice of the importance distribution $\eta$, that needs to be close enough to $\pi$.

@@

					\end{tikzpicture}
				\end{tcolorbox}
			\end{minipage}
			\caption{Flow chart of the Annealed Importance Sampling - Sequential Monte Carlo algorithm}
			\label{ASMCflowchart}
		\end{figure}

@@

	\paragraph{}
	Once the ASMC loop is over, we end up with a total of $n_{\rm particles}$ particles, which in principle should be close to annihilate our data, but whose coefficients may need some refinement. To deal with it, for each polynomial we run a fine tuning loop as described in sec.~\ref{sec:1runanalysis}.

@@

	\subsection{Sampling the manifold: gradient descent} \label{sec:graddes}
	To perform the gradient descent, we initialise randomly and uniformly points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range~$[-1,1]$, which will be crucial for the symbolic regression.\footnote{The rationale is that our symbolic regression evaluates the data points $\vec{x}^{(i)}$ using a candidate polynomial function $z$. When all data points lie within the interval $[-1, 1]$, the algorithm tends to favor high-degree polynomials artificially lowering the loss. Introducing data points with absolute values greater than 1 mitigates this bias.} As the example we are focussing on has a five-dimensional parameter space, we choose to generate $n_{\rm points}=10^5$ points, since we aim at populating all five directions and want approximately $\mathcal{O}(10)$ points per direction. As the solutions lie within the five dimensional space, its intrinsic dimension is less than or equal to 5. The value $10^5$ then serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

	We then perform a gradient descent on the points using TensorFlow’s automatic differentiation \cite{tensorflow2015-whitepaper}. The loss function is defined as
	\begin{equation} \label{eq:lossgraddes}
		\mathcal{L} = \sum_{i = 1}^{n_{\rm points}} \left|\left|\left.\nabla V(\vec{X}^{(i)})\,\right|_{\vec{y}^{\,(i)}=0}\right|\right|^2,
	\end{equation}
	% where $\vec{X}^{(i)} = (x^{(i)}_1,\dots,x^{(i)}_6,\tilde{x}^{(i)}_7,\tilde{x}^{(i)}_8,\tilde{x}^{(i)}_9,x^{(i)}_{10},\dots,x^{(i)}_{12},\tilde{x}^{(i)}_{13})$ and $\vec{y}^{\,(i)} = (x^{(i)}_3,x^{(i)}_5,x^{(i)}_6,\tilde{x}^{(i)}_7,\tilde{x}^{(i)}_9,x^{(i)}_{11},x^{(i)}_{12},\tilde{x}^{(i)}_{11})$
with the vectors $\vec{X}$ and $\vec{y}$ given in~\eqref{eq:defvecX} and~\eqref{eq:defvecy}. The exponents $(i)$ denote the data points, with $i\in\llbracket1,n_{\rm points}\rrbracket$. This way we only keep $(x^{(i)}_1,x^{(i)}_2,x^{(i)}_4,\tilde{x}^{(i)}_8,x^{(i)}_{10})$ alive after we have taken the derivative. As already mentioned above, we use the analytic formula for $\nabla V$, but we could have performed the gradient descent by fully using automatic differentiation.

@@

	The Adam optimizer was employed throughout the gradient descent procedure~\cite{kingma2017adammethodstochasticoptimization}. As an accelerator of convergence, we observed that periodically reinitialising the optimizer significantly improved the convergence rate. It was reinitialised at iterations {250, 500, 750}, with the learning rate $\alpha$ fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialised once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialisation was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs. The evolution of the loss function~\eqref{eq:lossgraddes}, along with the learning rate schedule, is shown in fig.~\ref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is rebooted. We also observe that, following the last few reinitialisations, the loss exhibits a small bump immediately after the restart. We interpret this behaviour as follows: the learning rate gets internally adjusted by the Adam optimizer during the descent, and may well be smaller than the instructions when the reset occurs. The learning rate then gets suddenly increased, and some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect of faster convergence after the reinitialization is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. This strategy introduces the risk of desampling certain regions in favour of others, introducing the risk to hide some flat directions in the potential. However, we believe that with enough points, it is very unlikely for a flat direction to be completely desampled. This method bears similarity to the concept of warm restarts introduced in ref.~\cite{loshchilov2017sgdrstochasticgradientdescent}, where the authors reset the learning rate to some value at each reset, without reinitialising the whole optimizer. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialised, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.

@@

	Upon completion, the loss function converges around $10^{-3}$. The distribution of the values of the gradient at each of the data points is plotted in fig.~\ref{fig:log_grad_V}. More than $93\%$ of the points converged to values of $\vert\nabla V\vert$ lower than $10^{-4}$, and more than $99\%$ of them exhibited gradient norms smaller than $10^{-3}$, yielding a satisfying sampling of the flat directions. As can be observed in fig.~\ref{fig:log_Vp4}, all data points converged to values close to $V(0)=-4$. Specifically, nearly all points attained $V(0)$ within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were of the order of $10^{-3}$ (not visible on the graph). As the origin $\vec{X}=0$ of the parameter space corresponds to the three-dimensional truncation of the round ${\rm AdS}_{3}\times S^{3}$ solution of half-maximal supergravity in six dimensions with $V(0)=-4$, this confirms that the data points lie within flat directions of the potential.

@@

	As a first visualisation, we present a tomography of the data in fig.~\ref{triangular_plot_1_2_4_8_10}. This figure shows all possible 2d projections of the data, along with the 1d histograms of each coordinate after gradient descent. Note that all directions appear to be well populated. There also seem to be non-trivial correlations in the data, see for example the $x_1/x_2$, $x_4/\tilde{x}_8$ or $x_4/x_{10}$ graphs. They could be genuine correlations, or may result from larger basins of attraction. Thanks to the subsequent analysis, we find that the structures in the $x_1/x_2$ plot are artefacts of the gradient descent, whereas the ones seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots reflect genuine features of the manifold.

	\subsection{Local analysis: extracting the dimension}
	Once the gradient descent has been completed and the flat directions sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we first perform some exploratory analyses to better understand the data. Specifically, we aim at determining the dimension of the manifold and whether it consists of a single connected component or multiple disjoint components (\textit{e.g.} two intersecting hyperplanes). To this end, we apply a local Principal Component Analysis (PCA).\footnote{Principal Component Analysis is a dimensionality reduction technique that transforms high-dimensional data into a lower-dimensional space while preserving the maximum amount of variance, see for example ref.~\cite{scikit-learn}.} % The method identifies orthogonal linear combinations of the original variables, called principal components, which are ordered by the proportion of total variance they explain. PCA operates by computing the eigenvalues and eigenvectors of the data covariance matrix, where the eigenvectors define the direction of maximum variance and the eigenvalues quantify the amount of variance captured along each direction. From a physicist point of view, as we are doing it locally, this gives access to information about the local tangent plane.}
	For each point, we identify its $k$ nearest neighbours and perform a PCA on that local neighbourhood. This procedure allows us to determine how many principal directions are needed to explain a given proportion $\epsilon$ of the data variance. In other words, it provides an estimate of the local dimensionality around each point, \textit{i.e.} the dimension of its local tangent space. We perform this analysis for several values of $k$, namely $k \in \{5, 10, 20, 50, 100\}$, and we fix $\epsilon = 0.99$. The results are presented in fig.~\ref{fig:local_pca}. We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned dimension 2. This can be attributed to the fact that if the genuine dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points. Additionally, for $k \geq 20$, we observe an increasing number of points being assigned dimensions 4 or 5. This behaviour can be explained by the loss of locality when the number of neighbours becomes too large: increasing $k$ results in a coarser approximation, and the algorithm may then incorporate points that are no longer truly local. This artificial enlargement of the neighbourhood can cause the estimated local dimensionality to rise. We thus infer that the manifold under investigation has an intrinsic dimension of 3.

@@

	\texttt{x\_10 = 1.414213551821*(x\_4-((x\_1/x\_2)-((x\_1/x\_2)/x\_8))}

>>>>

{\Large Machine Learning the Conformal Manifold of Holographic CFT$_{2}$’s}

@@

\begin{abstract}
	We investigate the structure of conformal manifolds around AdS$_3 \times S^3$ which lift from continuous flat directions in the scalar potential of gauged supergravity resulting from six-dimensional $\mathcal{N}=(1,1)$ supergravity.
	Our approach combines numerical exploration and symbolic inference. For the latter, we develop a symbolic regression algorithm based on Annealed Sequential Monte Carlo samplers, a combination of Annealed Importance Sampling and Sequential Monte Carlo samplers, well-suited to uncovering polynomial constraints in high-dimensional parameter spaces. 
	The algorithm reconstructs a set of polynomial relations that provides an explicit analytic parametrization of a new family of solutions. 
\end{abstract}

@@

Machine learning strategies have previously been applied to identify new isolated vacua in $\mathrm{SO}(8)$ supergravity \cite{Comsa:2019rcz,Berman:2022jqn}. More broadly, there has been increasing interest in applying machine learning and numerical techniques across various domains of high-energy physics. This includes, for instance, the characterisation of Calabi–Yau metrics and hypersurfaces \cite{Ashmore:2019wzb,Berman:2021mcw,Larfors:2022nep,Berglund:2022gvm,Jejjala:2020wcc,Douglas:2006rr,Larfors:2021pbb,He:2018jtw}, as well as broader efforts to explore the string theory landscape \cite{He:2017aed,Carifio:2017bov,Ruehle:2017mzq,Walden:2025cpf}. Additional applications include studies in CFT \cite{Chen:2020dxg}, investigations of the supergravity landscape \cite{Brady:2025zzi,Krishnan:2020sfg}, and explorations of the AdS/CFT correspondence \cite{Hashimoto:2018ftp}. More generally, machine learning has found utility in the study of string theory, geometry, and fundamental physics \cite{Ruehle:2020jrk,He:2023csq,Bao:2021auj}.

In this work, we demonstrate the viability of the aforementioned machine learning approach by applying it to a five-scalar subsector of a 13 scalar consistent truncation of six-dimensional non-chiral $\mathcal{N} = (1,1)$ supergravity on ${\rm AdS}_3 \times S^3$, or to type IIB supergravity on ${\rm AdS}_3 \times S^3 \times T^4$. Our methodology combines gradient descent sampling of the conformal manifold with a symbolic regression technique. There exists a large body of literature on symbolic regression, using methods from genetic programming \cite{koza1994genetic,virgolin2021improving,randall2022bingo,burlacu2019parsimony}, to deep learning \cite{petersen2019deep,kamienny2022end}, generative models \cite{valipour2021symbolicgpt}, diffusion models \cite{bastiani2025diffusion}, and equation learning with the nodes being symbolic operations \cite{2018arXiv180607259S}. A state-of-the-art algorithm is AIFeynman \cite{Udrescu:2019mnk}. It uses neural networks to identify structures in the dataset (such as translational symmetries, multiplicative separability, compositionality\dots) to recursively define simpler problems on which they can fit the solutions with polynomials and basic functions.  However, this method is slow, usually limited to low-dimensional spaces, and, by construction, can only fit one expression at a time. Alternatively, we develop here a symbolic regression algorithm based on an Annealed Sequential Monte Carlo sampler (ASMC)~\cite{zimmermann2021nested,syed2024optimisedannealedsequentialmonte} (combining the Annealed Importance Sampling~\cite{neal1998annealedimportancesampling} approach with a Sequential Monte Carlo sampler (SMC)~\cite{del2006sequential}), since this approach is much better suited to the study of geometric loci defined by intersecting polynomials in high-dimensional spaces.

@@

%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%
\section{Supergravity Setup} \label{sec:sugra}
Three-dimensional $\mathcal{N}=8$ (half-maximal) gauged supergravity is governed by the Lagrangian~\cite{Nicolai:2001ac,deWit:2003ja}
%
\begin{equation}	\label{eq: lagrangian_rephrased} 
	e^{-1}\mathcal{L}=R+\frac1{8}g^{\mu\nu}D_\mu M^{\bM\bN}D_\nu M_{\bM\bN}+e^{-1}\mathcal{L}_{\text{CS}}-V,
\end{equation}
%
which comprises an Einstein-Hilbert term $R$, a kinetic term for scalar fields parametrised by the matrix $M_{\bM\bN}$, a Chern-Simons contribution $\mathcal{L}_{\text{CS}}$, and a scalar potential $V$, see app.~\ref{app:sugra} for details on the definition of the terms. The scalar degrees of freedom parametrise the coset space
%
\begin{equation}	\label{eq: scalarcoset_rephrased}
	\frac{\text{SO}(8,4)}{\text{SO}(8)\times\text{SO}(4)},
\end{equation}
%
through the symmetric matrix $M_{\bK\bL}$.\footnote{In full generality, the coset space of half-maximal supergravity in three dimensions is $\text{SO}(8,p)/\big(\text{SO}(8)\times\text{SO}(p)\big)$. We consider here the case $p=4$ only.} With appropriate gauging (see eq.~\eqref{eq:embedtens}), this $\text{SO}(8,4)$ theory is a consistent truncation of six-dimensional ${\cal N}=(1,1)$ supergravity on $S^{3}$~\cite{Cvetic:2000dm,Deger:2014ofa,Hohm:2017wtr}. It features an ${\rm AdS}_{3}$ stationary point at the scalar origin ($M_{\bK\bL}=\delta_{\bK\bL}$, the ${\rm SO}(8,4)$ identity matrix), corresponding to an ${\rm AdS}_{3}\times S^{3}$ solution in 6d. In the following, we will be interested in exploring flat directions of the potential $V$ around this point, \textit{i.e.} stationary points that are continuously connected to the origin. In three dimensions, the flat directions constitute a family of ${\rm AdS}_{3}$ vacua sharing the same cosmological constant. The corresponding solutions in six dimensions are of the form ${\rm AdS}_{3}\times M^{3}$, with $M^{3}$ some deformation of the round sphere $S^{3}$, parametrised by the constant scalar vevs defining the 3d vacua.

@@

	To address these challenges, we formulate the search for annihilating polynomials as a determination of a probability density on the space of polynomials that stresses higher probabilities on polynomials $p$ that minimise the error
	\begin{equation}
		\sum_{i=1}^{n_{\rm points}}p\big(\vec{x}^{(i)}\big)^{2}.
	\end{equation}
	Polynomials sampled from this density would thus be good approximations to annihilating polynomials. Such a probability density is \textit{a priori} hard to find, but can be approximated using algorithms based on Sequential Important Sampling~\cite{Doucet2001,Liu2004} such as Sequential Monte Carlo samplers~\cite{del2006sequential}. They are inspired from Important Sampling (IS)~\cite{osti_4441935,RobertCasella2004}, a method to parametrise expectation values with respect to a target distribution $\pi$, that can be evaluated pointwise, but from which one does not know how to draw samples. IS uses another, typically simpler, distribution $\eta$ (called importance distribution) that one knows how to draw from. It is based on the observation that the expectation value
	\begin{equation}
		\mathbb{E}_{\pi}[f] = \int f(z)\pi(z)\ \d z,
	\end{equation}
	for some function $f$ with respect to $\pi$, can be computed from the expectation $\mathbb{E}_{\eta}[w f]$, now computed with respect to the distribution $\eta$, with unnormalised weights $w = \pi/\eta$. Therefore, one can estimate $\mathbb{E}_{\pi}[f]$ from a Monte Carlo method: we draw a sample $\{z^{(k)}\}_{k\in\llbracket 1, n_{\rm particles}\rrbracket}$ from $\eta$ and use the weighted average
	\begin{equation}
		\mathbb{E}_{\pi}[f] \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} w\big(z^{(k)}\big)\,f\big(z^{(k)}\big)\Bigg/\displaystyle \sum_{i=1}^{n_{\rm particles}} w\big(z^{(k)}\big).
	\end{equation}
	Stated differently, the target distribution $\pi$ can be approximated from a weighted collection of particles $\{z^{(k)},w^{(k)}\}_{k\in\llbracket 1, n_{\rm particles}\rrbracket}$ drawn from the importance distribution $\eta$ :
	\begin{equation}
		\pi(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)}\,\delta\big(z-z^{(k)}\big)\Bigg/\displaystyle \sum_{k=1}^{n_{\rm particles}} w^{(k)},
	\end{equation}
	where $w^{(k)} = w\big(z^{(k)}\big)$. The sample admits $\pi$ as marginal distribution asymptotically when $n_{\rm particles}\rightarrow+\infty$. The accuracy of this estimation depends on the size of the sample and, crucially, on the choice of the importance distribution $\eta$, that needs to be close enough to $\pi$.

@@

					\end{tikzpicture}
				\end{tcolorbox}
			\end{minipage}
			\caption{Flow chart of the Annealed Sequential Monte Carlo sampler algorithm.}
			\label{ASMCflowchart}
		\end{figure}

@@

	\paragraph{}
	Once the ASMC loop is over, we end up with a total of $n_{\rm particles}$ polynomials, which in principle should be close to annihilate our data, but whose coefficients may need some refinement. To deal with it, for each polynomial we run a fine tuning loop as described in sec.~\ref{sec:1runanalysis}.

@@

	\subsection{Sampling the manifold: gradient descent} \label{sec:graddes}
	To perform the gradient descent, we initialise randomly and uniformly points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range~$[-1,1]$, which will be crucial for the symbolic regression.\footnote{The rationale is that our symbolic regression evaluates the data points $\vec{x}^{(i)}$ using a candidate polynomial function~$z$. When all data points lie within the interval $[-1, 1]$, the algorithm tends to favor high-degree polynomials artificially lowering the loss. Introducing data points with absolute values greater than 1 mitigates this bias.} As the example we are focussing on has a five-dimensional parameter space, we choose to generate $n_{\rm points}=10^5$ points, since we aim at populating all five directions and want approximately $\mathcal{O}(10)$ points per direction. As the solutions lie within the five dimensional space, its intrinsic dimension is less than or equal to 5. The value $10^5$ then serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

	We then perform a gradient descent on the points using TensorFlow’s automatic differentiation \cite{tensorflow2015-whitepaper}. The loss function is defined as
	\begin{equation} \label{eq:lossgraddes}
		\mathcal{L} = \sum_{i = 1}^{n_{\rm points}} \left|\left|\left.\nabla V(\vec{X}^{(i)})\,\right|_{\vec{y}^{\,(i)}=0}\right|\right|^2,
	\end{equation}
	% where $\vec{X}^{(i)} = (x^{(i)}_1,\dots,x^{(i)}_6,\tilde{x}^{(i)}_7,\tilde{x}^{(i)}_8,\tilde{x}^{(i)}_9,x^{(i)}_{10},\dots,x^{(i)}_{12},\tilde{x}^{(i)}_{13})$ and $\vec{y}^{\,(i)} = (x^{(i)}_3,x^{(i)}_5,x^{(i)}_6,\tilde{x}^{(i)}_7,\tilde{x}^{(i)}_9,x^{(i)}_{11},x^{(i)}_{12},\tilde{x}^{(i)}_{11})$
with the vectors $\vec{X}$ and $\vec{y}$ given in~\eqref{eq:defvecX} and~\eqref{eq:defvecy}. The exponents $(i)$ denote the data points, with $i\in\llbracket1,n_{\rm points}\rrbracket$. This way we only keep $\vec{x}^{(i)} = (x^{(i)}_1,x^{(i)}_2,x^{(i)}_4,\tilde{x}^{(i)}_8,x^{(i)}_{10})$ alive after we have taken the derivative. As already mentioned above, we use the analytic formula for $\nabla V$, but we could have performed the gradient descent by fully using automatic differentiation.

@@

	The Adam optimizer was employed throughout the gradient descent procedure~\cite{kingma2017adammethodstochasticoptimization}. As an accelerator of convergence, we observed that periodically reinitialising the optimizer significantly improved the convergence rate. It was reinitialised at iterations {250, 500, 750}, with the learning rate $\alpha$ fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialised once more, this time with a reduced learning rate of~$10^{-3}$. A final reinitialisation was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs. The evolution of the loss function~\eqref{eq:lossgraddes}, along with the learning rate schedule, is shown in fig.~\ref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is rebooted. We also observe that, following the last few reinitialisations, the loss exhibits a small bump immediately after the restart. We interpret this behaviour as follows: the learning rate gets internally adjusted by the Adam optimizer during the descent, and may well be smaller than the instructions when the reset occurs. The learning rate then gets suddenly increased, and some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect of faster convergence after the reinitialization is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. This strategy introduces the risk of desampling certain regions in favour of others, introducing the risk to hide some flat directions in the potential. However, we believe that with enough points, it is very unlikely for a flat direction to be completely desampled. This method bears similarity to the concept of warm restarts introduced in ref.~\cite{loshchilov2017sgdrstochasticgradientdescent}, where the authors reset the learning rate to some value at each reset, without reinitialising the whole optimizer. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialised, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.

@@

	Upon completion, the loss function converges around $10^{-3}$. The distribution of the values of the gradient at each of the data points is plotted in fig.~\ref{fig:log_grad_V}. More than $93\%$ of the points converged to values of $\vert\nabla V\vert$ lower than $10^{-4}$, and more than $99\%$ of them exhibited gradient norms smaller than~$10^{-3}$, yielding a satisfying sampling of the flat directions. As can be observed in fig.~\ref{fig:log_Vp4}, all data points converged to values close to $V(0)=-4$. Specifically, nearly all points attained $V(0)$ within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were of the order of $10^{-3}$ (not visible on the graph). As the origin $\vec{X}=0$ of the parameter space corresponds to the three-dimensional truncation of the round ${\rm AdS}_{3}\times S^{3}$ solution of half-maximal supergravity in six dimensions with $V(0)=-4$, this confirms that the data points lie within flat directions of the potential.

@@

	As a first visualisation, we present a tomography of the data in fig.~\ref{triangular_plot_1_2_4_8_10}. This figure shows all possible 2d projections of the data, along with the 1d histograms of each coordinate after gradient descent. Note that all directions appear to be well populated. There also seem to be non-trivial correlations in the data, see for example the $x_1/x_2$, $x_4/\tilde{x}_8$ or $x_4/x_{10}$ graphs. They could be genuine correlations, or may result from larger basins of attraction. Thanks to the analysis of sec.~\ref{sec:localanalysis} and~\ref{sec:AIS-SMC_application}, we find that the structures in the $x_1/x_2$ plot are artefacts of the gradient descent, whereas the ones seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots reflect genuine features of the manifold.

	\subsection{Local analysis: extracting the dimension} \label{sec:localanalysis}
	Once the gradient descent has been completed and the flat directions sampled, the next step is to identify the structure of the underlying manifold. Our goal is to eventually obtain an analytical expression, not just a numerical description. Before applying symbolic regression to search for such an expression, we first perform some exploratory analyses to better understand the data. Specifically, we aim at determining the dimension of the manifold and whether it consists of a single connected component or multiple disjoint components (\textit{e.g.} two intersecting hyperplanes). To this end, we apply a local Principal Component Analysis (PCA).\footnote{Principal Component Analysis is a dimensionality reduction technique that transforms high-dimensional data into a lower-dimensional space while preserving the maximum amount of variance, see for example ref.~\cite{scikit-learn}.} % The method identifies orthogonal linear combinations of the original variables, called principal components, which are ordered by the proportion of total variance they explain. PCA operates by computing the eigenvalues and eigenvectors of the data covariance matrix, where the eigenvectors define the direction of maximum variance and the eigenvalues quantify the amount of variance captured along each direction. From a physicist point of view, as we are doing it locally, this gives access to information about the local tangent plane.}
	For each point, we identify its $k$ nearest neighbours and perform a PCA on that local neighbourhood. This procedure allows us to determine how many principal directions are needed to explain a given proportion $\epsilon$ of the data variance. In other words, it provides an estimate of the local dimensionality around each point, \textit{i.e.} the dimension of its local tangent space. We perform this analysis for several values of $k$, namely $k \in \{5, 10, 20, 50, 100\}$, and we fix $\epsilon = 0.99$. The results are presented in fig.~\ref{fig:local_pca}. We observe that for every choice of $k$, there is a prominent peak at $d = 3$, suggesting that the underlying manifold is three-dimensional. For $k = 5$, a noticeable fraction of points are assigned dimension 2. This can be attributed to the fact that if the genuine dimension is 3, then selecting only 5 neighbours may not sufficiently populate all three directions, leading the algorithm to underestimate the dimensionality for a fraction of the points. Additionally, for $k \geq 20$, we observe an increasing number of points being assigned dimensions 4 or 5. This behaviour can be explained by the loss of locality when the number of neighbours becomes too large: increasing $k$ results in a coarser approximation, and the algorithm may then incorporate points that are no longer truly local. This artificial enlargement of the neighbourhood can cause the estimated local dimensionality to rise. We thus infer that the manifold under investigation has an intrinsic dimension of 3.

@@

	\texttt{x\_10 = 1.414213551821*(x\_4-((x\_1/x\_2)-((x\_1/x\_2)/x\_8))},
